Missing Data: Current Practice in Football Research and Recommendations for

### 2020-11-09 paper/paper.Rmd
edit formatting for tex.

---
title: 'Missing Data: Current Practice in Football Research and Recommendations for
  Improvement'
bibliography: references.bib
csl: sport-medicine.csl
output:
  pdf_document:
    toc: no
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
  bookdown::pdf_document2:
    toc: no
  word_document:
    toc: yes
authors: David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(out.width = "75%",
                      dev = "png",
                      fig.align = "center",
                      echo = FALSE)
```

@@

**Authors**

David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)

**Affiliations**

1: Griffith University, Menzies Health Institute Queensland, The Hopkins Centre, Brisbane, Australia.
2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.
3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.
4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.
5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.

**Twitter handles**

elborgo9, roberto_boberto, nj_tierney

**Conflict of interest**

@@

**Corresponding author**

Nicholas J Tierney Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia. Address: Email: [nicholas.tierney\@monash.edu](mailto:nicholas.tierney@monash.edu){.email}

**Data availability**

The code to reproduce the results and plots in the paper can be accessed at the following GitHub repository: [https://github.com/SciBorgo/Missing-data.git](https://github.com/SciBorgo/Missing-data.git).

@@

# Methods and materials

@@

## Practices to avoid

@@

## Common causes of missing values

@@

@@

## Visualising missing values

We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017]. The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.

```{r missing-overview, fig.cap = "**Overviews of missing values in simulated injury data in Australian Football matches for August, 2020. The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match. Researchers might be interested in investigating the relationship between number of sprints and total sprint distance with incidence of injury. Panel A shows... Panel B shows ... Panel C shows ... Panel D shows ...", fig.height = 10}

@@

gg_miss_scatter <- ggplot(simulated_afl_data,
       aes(x = sprint_dist,
           y = sprints)) +
  geom_miss_point(size = 2) + 
  theme_minimal() +
  #theme(text = element_text(size = 14)) +
  labs(x = "Total sprint distance (m)", 
       y = "Number of sprints (count)", 
       colour = "",
       tag = "(C)")

@@

```{r patch-2, fig.height = 10}
(aq_vis_miss + aq_fct) / (gg_miss_scatter + aq_miss_var) & 
  theme(aspect.ratio = 1,
        legend.position = "bottom")
```

To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].

## Understanding imputation

>>>>

---
title: 'Missing Data: Current Practice in Football Research and Recommendations for
  Improvement'

bibliography: references.bib
csl: sport-medicine.csl
output:
  pdf_document:
    toc: no
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
  bookdown::pdf_document2:
    toc: no
  word_document:
    toc: yes
authors: David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(out.width = "75%",
                      dev = "png",
                      fig.align = "center",
                      echo = FALSE,
                      dpi = 600)
```

@@

\begin{center}
David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
\end{center}
\vspace{5mm}
**Affiliations**

1: Griffith University, Menzies Health Institute Queensland, The Hopkins Centre, Brisbane, Australia.\hfill \break
2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.\hfill \break
3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.\hfill \break
4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.\hfill \break
5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.

\vspace{5mm}
**Citation:** Borg DN, Nguyen R, and Tierney NJ. (2020). Missing Data: Current Practice in Football Research and Recommendations for Improvement.

\vspace{5mm}
**Twitter:** elborgo9, roberto_boberto, nj_tierney

\vspace{5mm}
**Corresponding Author**

Nicholas J Tierney\hfill \break
Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.\hfill \break
Email: [nicholas.tierney\@monash.edu](mailto:nicholas.tierney@monash.edu){.email}

\vspace{5mm}
**Conflict of Interest**

@@

\vspace{5mm}
**Funding**

No funding was awarded for this project.

\vspace{5mm}
**Data Availability**

The code to reproduce the results and plots in the paper can be accessed at the GitHub repository: [https://github.com/SciBorgo/Missing-data.git](https://github.com/SciBorgo/Missing-data.git).

@@

# Methods and Materials

@@

## Practices to Avoid

@@

## Common Causes of Missing Values

@@

\vspace{5mm}

@@

## Visualising Missing Values

```{r}

@@

gg_miss_scatter <- ggplot(simulated_afl_data,
       aes(x = sprint_dist,
           y = sprints)) +
  geom_miss_point(size = 2, alpha = 0.15) + 
  theme_minimal() +
  #theme(text = element_text(size = 14)) +
  labs(x = "Total sprint distance (m)", 
       y = "Number of sprints (count)", 
       colour = "",
       tag = "(C)")

@@

We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017]. The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in an simulated injury dataset. To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].

```{r missing-overview,  fig.cap = "Overviews of missing values in simulated injury data from Australian Football matches played in August, 2020. The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match. Panel A shows the location of missing (black) and complete (grey) cases, and the percentage of missing data in each variable and overall. Panel B shows the percentage of missings for each team, in each variable. Panel C shows the relationship between two variables in the dataset, including missing values (red) which are shifted, so that they can be visualised on the same axes as the regular values. Panel D shows the number of missing values in each variable", fig.height = 7.75}

(aq_vis_miss + aq_fct)/(gg_miss_scatter + aq_miss_var) & 
  theme(aspect.ratio = 1,
        legend.position = "bottom")
```

## Understanding Imputation
